---
title: "US Politics Part II"
author: "Christopher Marks"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{US Politics Part II}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# PART II: Update Data and Analyze

This is the second part of a two-part vignette designed to collect and analyze Twitter posts from members of the US Congress and the US President. This part of the vignette carries out the database initialization and initial status collections. This part of the vignette consists of two sub-parts: (a) updating the twitter database with new posts, and (b) analysis.  These two sub-parts can be run together or independently, as often as required.

**Very Important Note:** As with the first part of this vignette, Twitter data collection can take a very long time.  In this part of the vignette, collection is carried out using the `update_user_timelines` function.  If the Twitter database has been kept up to date, i.e., if this function has been called recently on the Twitter data connection, then there will not be many new statuses and the method will run in a few minutes.  However, if this function has not been called for a longer period of time, and the database consists of all members of congress, then there will be more new statuses to collect and the function will take longer to complete.


# Initialization

We begin by allocating additional memory for RJava to support sentiment analysis, loading the `twinfoR` namespace, establishing the Twitter database connection to the sqlite database created in Part I of this vignette, and loading the authentication vector.  See (and execute) the `us_politics_1` vignette if you haven't already done so.

This code assumes the `auth.vector` from Part I is saved.  If it isn't, reauthenticate using the `authorize_IT()` function as done in Part I.

```{r eval = TRUE}
options(java.parameters = "-Xmx1024m")
library(twinfoR)
setwd("~/USPol")
con <- twitter_database("pol.sqlite")
load("auth_vector.RData") 
```

# Update Database with New Statuses

This function should take considerably less time to run than the initial call, because it only collects new statuses.  Again, we divert output to the NULL device.

```{r eval=FALSE, message=FALSE}
# sink("/dev/null") # Recommended to suppress Rsentiment output, but
## might not work on all platforms.
update_user_timelines(
  con,
  calc.Rsentiment = TRUE,
  calc.syu = TRUE
)
# sink()
```
```{r}
# Summarize database
summarize_database(con)
```

# Analysis

This section goes quickly through the analysis capability available in the `twinfoR` package.

## Top hashtags from democrats and republicans in the last week

```{r}
tophashtags.dem <- top_hashtags(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'democrat'"
)
names(tophashtags.dem) = c("Dem.Hashtag.Lower","Count","Dem.Hashtag")
tophashtags.dem$Dem.Hashtag <- paste("#",tophashtags.dem$Dem.Hashtag,sep="")

tophashtags.rep <- top_hashtags(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican'"
)
names(tophashtags.rep) = c("Rep.Hashtag.Lower","Count","Rep.Hashtag")
tophashtags.rep$Rep.Hashtag <- paste("#",tophashtags.rep$Rep.Hashtag,sep="")

knitr::kable(
  list(
    tophashtags.dem[,c("Dem.Hashtag","Count")],
    tophashtags.rep[,c("Rep.Hashtag","Count")]
  ),
  caption = "Top Hashtags Used by Democrats and Republicans"
)
```

## Top usermentions from democrats and republicans in the last week

```{r}
topusermentions.dem <- top_usermentions(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'democrat'"
)

topusermentions.rep <- top_usermentions(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican'"
)


knitr::kable(
  data.frame(
    Screen_Name = paste("@",topusermentions.dem$user_mention_screen_name,sep=""),
    Name = topusermentions.dem$user_mention_name,
    Mentions = topusermentions.dem$n,
    stringsAsFactors = FALSE
  ),
  caption = "Democrat User Mentions"
)
knitr::kable(
  data.frame(
    Screen_Name = paste("@",topusermentions.rep$user_mention_screen_name,sep=""),
    Name = topusermentions.rep$user_mention_name,
    Mentions = topusermentions.rep$n,
    stringsAsFactors = FALSE
  ),
  caption = "Republican User Mentions"
)
```

## Top URLs posted by democrats and republicans in the last week

```{r}
topurls.dem <- top_urls(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'democrat'"
)

topurls.rep <- top_urls(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican'"
)


knitr::kable(
  list(
    Democrat = data.frame(
      Dem.URL = topurls.dem$extended_url,
      Count = topurls.dem$n,
      stringsAsFactors = FALSE
    ),
    Republican = data.frame(
      Rep.URL = topurls.rep$extended_url,
      Count = topurls.rep$n,
      stringsAsFactors = FALSE
    )
  ),
  caption = "Top Democrat and Republican URLs"
)
```

### Top media from democrats and republicans in the last week

```{r fig.width = 4, fig.height = 4}
topmedia.dem <- top_media(
  con,
  start.date = Sys.Date()-14,
  where.criteria = "query_users.party = 'democrat'",
  save.to.file = FALSE
)

topmedia.rep <- top_media(
  con,
  start.date = Sys.Date()-14,
  where.criteria = "query_users.party = 'republican'",
  save.to.file = FALSE
)

knitr::kable(
  list(
    data.frame(
      Dem.image_url = topmedia.dem$media_url,
      count = topmedia.dem$n,
      stringsAsFactors = FALSE
    ),
    data.frame(
      Rep.image_url = topmedia.rep$media_url,
      count = topmedia.rep$n,
      stringsAsFactors = FALSE
    )
  ),
  caption = "Top Media Posted by Democrats and Republicans"
)

## Top Dem Image (Not Run)

# status_media(
#   topmedia.dem[1,],
#   con, 
#   update.media.b64 = FALSE,
#   display.image = TRUE
# )

## Top Rep Image (Not Run)

# status_media(
#   topmedia.rep[1,],
#   con,
#   update.media.b64 = FALSE,
#   display.image = TRUE
# )
```

```{r echo = FALSE}
dem.split <- strsplit(topmedia.dem$media_url[1],".",fixed=TRUE)[[1]]
dem.ext <- dem.split[length(dem.split)]
if(tolower(dem.ext) %in% c("jpg","jpeg")){
  dem.media.type <- "jpeg"
} else if(tolower(dem.ext) %in% c("tif","tiff")){
  dem.media.type <- "tiff"
} else {
  dem.media.type <- "png"
}
rep.split <- strsplit(topmedia.rep$media_url[1],".",fixed=TRUE)[[1]]
rep.ext <- rep.split[length(rep.split)]
if(tolower(rep.ext) %in% c("jpg","jpeg")){
  rep.media.type <- "jpeg"
} else if(tolower(rep.ext) %in% c("tif","tiff")){
  rep.media.type <- "tiff"
} else {
  rep.media.type <- "png"
}
dem.b64 <- DBI::dbGetQuery(
  con,
  sprintf(
    "SELECT img_b64 FROM media WHERE media_id = '%s' LIMIT 1;",
    topmedia.dem$media_id[1]
  )
)$img_b64[1]
rep.b64 <- DBI::dbGetQuery(
  con,
  sprintf(
    "SELECT img_b64 FROM media WHERE media_id = '%s' LIMIT 1;",
    topmedia.rep$media_id[1]
  )
)$img_b64[1]
knitr::asis_output(
  htmltools::htmlPreserve(
    paste(
      "\n",
      "<div>\n",
      "<table>\n",
      "<tr>\n",
      "<td align=\"center\">\n",
      sprintf(
        "<img width=\"420px\" src=\"data:image/%s;base64,%s\" alt=\"Dem Image\"/>\n",
        dem.media.type,
        dem.b64
      ),
      "</td>\n",
      "<td align=\"center\">\n",
      sprintf(
        "<img width=\"420px\" src=\"data:image/%s;base64,%s\" alt=\"Rep Image\"/>\n",
        rep.media.type,
        rep.b64
      ),
      "</td>\n",
      "</tr>\n",
      "<tr>\n",
      "<td align = \"center\">\n",
      "Top Democrat Image\n",
      "</td>",
      "<td align = \"center\">\n",
      "Top Republican Image\n",
      "</td>\n",
      "</tr>\n",
      "</table>\n",
      "</div>\n"
    )
  )
)
```

The `top_media` and `status_media` have many options for saving and rendering images.  See documentation for details.

### Top tweeters democrats and republicans in the last week

```{r}
toptweeters.dem <- top_tweeters(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'democrat'"
)

toptweeters.rep <- top_tweeters(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican'"
)

toptweeters.dem$description <- gsub("\n"," -- ",toptweeters.dem$description)
toptweeters.dem$location <- gsub("\n"," -- ",toptweeters.dem$location)
toptweeters.rep$description <- gsub("\n"," -- ",toptweeters.rep$description)
toptweeters.rep$location <- gsub("\n"," -- ",toptweeters.rep$location)
knitr::kable(
  data.frame(
    Name = toptweeters.dem$name,
    Location = toptweeters.dem$location,
    Description = toptweeters.dem$description,
    Tweets = toptweeters.dem$n,
    stringsAsFactors = FALSE
  ),
  caption = "Top Democrat Tweeters"
)
knitr::kable(
  data.frame(
    Name = toptweeters.rep$name,
    Location = toptweeters.rep$location,
    Description = toptweeters.rep$description,
    Tweets = toptweeters.rep$n,
    stringsAsFactors = FALSE
  ),
  caption = "Top Republican Tweeters"
)
```

### Most liked tweets democrats and republicans in the last week

```{r}
mostliked.dem <- most_liked(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'democrat'"
)

mostliked.rep <- most_liked(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican'"
)

# For nice Markdown rendering
mostliked.dem$description <- gsub("\n"," -- ",mostliked.dem$description)
mostliked.dem$location <- gsub("\n"," -- ",mostliked.dem$location)
mostliked.dem$text <- gsub("\n"," -- ",mostliked.dem$text)
mostliked.dem$name <- gsub("\n"," -- ",mostliked.dem$name)
mostliked.rep$description <- gsub("\n"," -- ",mostliked.rep$description)
mostliked.rep$location <- gsub("\n"," -- ",mostliked.rep$location)
mostliked.rep$text <- gsub("\n"," -- ",mostliked.rep$text)
mostliked.rep$name <- gsub("\n"," -- ",mostliked.rep$name)
knitr::kable(
  data.frame(
    Name = mostliked.dem$name,
    Created = mostliked.dem$created_at,
    Likes = mostliked.dem$favorite_count,
    Text = mostliked.dem$text,
    stringsAsFactors = FALSE
  ),
  caption = "Most Liked Democrat Tweets"
)
knitr::kable(
  data.frame(
    Name = mostliked.rep$name,
    Created = mostliked.rep$created_at,
    Likes = mostliked.rep$favorite_count,
    Text = mostliked.rep$text,
    stringsAsFactors = FALSE
  ),
  caption = "Most Liked Republican Tweets"
)
```


### Most retweeted tweets democrats and republicans in the last week

```{r}
mostretweeted.dem <- most_retweeted(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'democrat'"
)

mostretweeted.rep <- most_retweeted(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican'"
)

# For nice Markdown rendering
mostretweeted.dem$description <- gsub("\n"," -- ",mostretweeted.dem$description)
mostretweeted.dem$location <- gsub("\n"," -- ",mostretweeted.dem$location)
mostretweeted.dem$text <- gsub("\n"," -- ",mostretweeted.dem$text)
mostretweeted.dem$name <- gsub("\n"," -- ",mostretweeted.dem$name)
mostretweeted.rep$description <- gsub("\n"," -- ",mostretweeted.rep$description)
mostretweeted.rep$location <- gsub("\n"," -- ",mostretweeted.rep$location)
mostretweeted.rep$text <- gsub("\n"," -- ",mostretweeted.rep$text)
mostretweeted.rep$name <- gsub("\n"," -- ",mostretweeted.rep$name)
knitr::kable(
  data.frame(
    Name = mostretweeted.dem$name,
    Created = mostretweeted.dem$created_at,
    Retweet_Count = mostretweeted.dem$retweet_count,
    Text = mostretweeted.dem$text,
    stringsAsFactors = FALSE
  ),
  caption = "Most Retweeted Democrat Tweets"
)
knitr::kable(
  data.frame(
    Name = mostretweeted.rep$name,
    Created = mostretweeted.rep$created_at,
    Retweet_Count = mostretweeted.rep$retweet_count,
    Text = mostretweeted.rep$text,
    stringsAsFactors = FALSE
  ),
  caption = "Most Retweeted Republican Tweets"
)
```


## Most popular retweet in sample

```{r}
mostpopRT.dem <- most_popular_RT_in_sample(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'democrat'"
)

mostpopRT.rep <- most_popular_RT_in_sample(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican'"
)

# For nice Markdown rendering
mostpopRT.dem$description <- gsub("\n"," -- ",mostpopRT.dem$description)
mostpopRT.dem$location <- gsub("\n"," -- ",mostpopRT.dem$location)
mostpopRT.dem$text <- gsub("\n"," -- ",mostpopRT.dem$text)
mostpopRT.dem$name <- gsub("\n"," -- ",mostpopRT.dem$name)
mostpopRT.rep$description <- gsub("\n"," -- ",mostpopRT.rep$description)
mostpopRT.rep$location <- gsub("\n"," -- ",mostpopRT.rep$location)
mostpopRT.rep$text <- gsub("\n"," -- ",mostpopRT.rep$text)
mostpopRT.rep$name <- gsub("\n"," -- ",mostpopRT.rep$name)
knitr::kable(
  data.frame(
    Name = mostpopRT.dem$name,
    Created = mostpopRT.dem$created_at,
    Dem_Retweets = mostpopRT.dem$n,
    Text = mostpopRT.dem$text,
    stringsAsFactors = FALSE
  ),
  caption = "Most Popular Retweets Among Democrats"
)
knitr::kable(
  data.frame(
    Name = mostpopRT.rep$name,
    Created = mostpopRT.rep$created_at,
    Rep_Retweets = mostpopRT.rep$n,
    Text = mostpopRT.rep$text,
    stringsAsFactors = FALSE
  ),
  caption = "Most Popular Retweets Among Republicans"
)
```


### Most reach 

Note: this metric is of dubious value.  See documentation.

```{r}
mostreach.dem <- most_reach(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'democrat'"
)

mostreach.rep <- most_reach(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican'"
)

# For nice Markdown rendering
mostreach.dem$description <- gsub("\n"," -- ",mostreach.dem$description)
mostreach.dem$location <- gsub("\n"," -- ",mostreach.dem$location)
mostreach.dem$text <- gsub("\n"," -- ",mostreach.dem$text)
mostreach.dem$name <- gsub("\n"," -- ",mostreach.dem$name)
mostreach.rep$description <- gsub("\n"," -- ",mostreach.rep$description)
mostreach.rep$location <- gsub("\n"," -- ",mostreach.rep$location)
mostreach.rep$text <- gsub("\n"," -- ",mostreach.rep$text)
mostreach.rep$name <- gsub("\n"," -- ",mostreach.rep$name)
knitr::kable(
  data.frame(
    Name = mostreach.dem$name,
    Created = mostreach.dem$created_at,
    Reach = mostreach.dem$n,
    Text = mostreach.dem$text,
    stringsAsFactors = FALSE
  ),
  caption = "Democrat Tweets with Most 'Reach'"
)
knitr::kable(
  data.frame(
    Name = mostreach.rep$name,
    Created = mostreach.rep$created_at,
    Reach = mostreach.rep$n,
    Text = mostreach.rep$text,
    stringsAsFactors = FALSE
  ),
  caption = "Republican Tweets with Most 'Reach'"
)
```


### Word clouds 

```{r fig.width=7,fig.height=7,message=FALSE,warning=FALSE,results="hold"}
wordcloud_plot(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'democrat'",
  caption = "Democrats"
)

wordcloud_plot(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican'",
  caption = "Republicans"
)
```

### Sentiment plots

Note these are saved to files.

```{r fig.width=3,fig.height=3}
sentiment_plots(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'democrat'",
  file.name.prefix = "Dem",
  caption = "Democrats"
)

sentiment_plots(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican'",
  file.name.prefix = "Rep",
  caption = "Republicans"
)
```

```{r echo=FALSE, fig.width=3, fig.height=3}
knitr::asis_output(
  htmltools::htmlPreserve(
    paste(
      "\n",
      "<div>\n",
      "<table>\n",
      "<tr>\n",
      "<td align=\"center\">\n",
      sprintf(
        "<img width=\"420px\" src=\"data:image/png;base64,%s\" alt=\"Dem Image\"/>\n",
        base64enc::base64encode("Dem_sentiment-barplot.png")
      ),
      "</td>\n",
      "<td align=\"center\">\n",
      sprintf(
        "<img width=\"420px\" src=\"data:image/png;base64,%s\" alt=\"Rep Image\"/>\n",
        base64enc::base64encode("Rep_sentiment-barplot.png")
      ),
      "</td>\n",
      "</tr>\n",
      "<tr>\n",
      "<td align = \"center\">\n",
      "Democrat Absolute Sentiment\n",
      "</td>",
      "<td align = \"center\">\n",
      "Republican Absolute Sentiment\n",
      "</td>\n",
      "</tr>\n",
      "<tr style=\"padding-top:30px\">\n",
      "<td align=\"center\">\n",
      sprintf(
        "<img width=\"420px\" src=\"data:image/png;base64,%s\" alt=\"Dem Image\"/>\n",
        base64enc::base64encode("Dem_sentiment-lineplot.png")
      ),
      "</td>\n",
      "<td align=\"center\">\n",
      sprintf(
        "<img width=\"420px\" src=\"data:image/png;base64,%s\" alt=\"Rep Image\"/>\n",
        base64enc::base64encode("Rep_sentiment-lineplot.png")
      ),
      "</td>\n",
      "</tr>\n",
      "<tr>\n",
      "<td align = \"center\">\n",
      "Democrat Relative Sentiment\n",
      "</td>",
      "<td align = \"center\">\n",
      "Republican Relative Sentiment\n",
      "</td>\n",
      "</tr>\n",
      "</table>\n",
      "</div>\n"
    )
  )
)
```

### Tweet timeplot

```{r fig.width=5, fig.height=5}
timeplot(
  con,
  start.date = Sys.Date()-7,
  where.criteria = "query_users.party = 'republican' OR query_users.party = 'democrat'",
  group.column = "query_users.party"
)
```


### Friend network analysis and some nice network plots


```{r}
# Import igraph
library(igraph)

# Custom database queries and graph creation.
edgelist.query <- "SELECT query_users_fo.screen_name as follower, query_users_fr.screen_name as friend, query_users_fo.party as follower_party, query_users_fr.party as friend_party FROM query_users as query_users_fo JOIN followers ON query_users_fo.user_id = followers.follower_id JOIN query_users AS query_users_fr ON followers.friend_id = query_users_fr.user_id;"
edgelist.df <- DBI::dbGetQuery(con,edgelist.query)
g <- graph_from_data_frame(edgelist.df)
vertices.df <- data.frame(index=as.integer(V(g)), screen_name = names(V(g)),stringsAsFactors = FALSE)
vertex.query <- sprintf("SELECT query_users.screen_name,query_users.party,query_users.body,user.name,user.location,user.description,user.followers_count,user.statuses_count FROM query_users JOIN user ON query_users.user_id = user.id WHERE query_users.screen_name IN (%s);",paste("'",vertices.df$screen_name,"'",sep="",collapse=","))
vertex.query.df <- DBI::dbGetQuery(con,vertex.query)
vertex.query.df <- vertex.query.df[match(vertices.df$screen_name,vertex.query.df$screen_name),]

bluered <- function(party){
  if(tolower(party)=="democrat"){
    return("blue")
  } else if(tolower(party)=="republican"){
    return("red")
  } else {
    return("black")
  }
}

vertex.query.df$color <- sapply(vertex.query.df$party,bluered)
```

#### Eigenvector Centrality

 ```{r}
ec <- centr_eigen(g,directed=FALSE)
df <- data.frame(
  Screen_Name = paste(
    "@",
    vertex.query.df$screen_name,
    sep=""
  ),
  Eig_Centrality = as.numeric(ec$vector)
)
df <- df[order(df[,2],decreasing=TRUE),]
knitr::kable(
  df[1:10,],
  row.names = FALSE,
  caption = "Friend Network Eigenvector Centrality"
)
```

#### Betweenness centrality

```{r}
bc <- centr_betw(g,directed=FALSE)
df <- data.frame(
  Screen_Name = paste(
    "@",
    vertex.query.df$screen_name,
    sep=""
  ),
  Bet_Centrality = as.numeric(bc$res)
)
df <- df[order(df[,2],decreasing=TRUE),]
knitr::kable(
  df[1:10,],
  row.names = FALSE,
  caption = "Friend Network Betweenness Centrality"
)
```

#### Clustering (label propagation)

```{r}
clust.labelprop <- cluster_label_prop(as.undirected(g))
clust.labelprop
```

#### Visualize

```{r fig.width=7,fig.height=7}
verts <- data.frame(
    name = vertex.query.df$screen_name,
    label = vertex.query.df$screen_name,
    title = vertex.query.df$name,
    color = vertex.query.df$color,
    stringsAsFactors = FALSE
)

for(n in names(verts)){
  g <- set_vertex_attr(g,n,value=verts[,n])
}

plot(g,vertex.color=vertex_attr(g,"color"),vertex.size=4,vertex.label=NA)
```


#### Interactive visualization with sigmajs

```{r eval = FALSE, fig.width=7,fig.height=7}
# Not run
library(sigmajs)

s <- sigmajs()
s <- sg_from_igraph(s,g)
s <- sg_settings(s,defaultEdgeColor="#AAA", edgeColor="default",minNodeSize=1,maxNodeSize=4)
s
```

#### Interactive visualization with visNetwork

```{r eval = FALSE, fig.width=7,fig.height=7}
# Not Run
library(visNetwork)
v <- visIgraph(g)
v <- visEdges(
  v,
  color="gray"
)
v <- visNodes(
  v,
  size = 10
)
v
```




### User mention network analysis

This is essentially the same analysis as above, but using the user mention network.

```{r fig.width=7, fig.height=7}
usermention.query <- sprintf("SELECT query_users_mentioner.screen_name as mentioner,query_users_mentioned.screen_name as mentioned, status.id as status_id, status.text as status_text, status.created_at as created_at FROM query_users as query_users_mentioner JOIN status ON query_users_mentioner.user_id = status.user_id JOIN user_mention ON status.id = user_mention.status_id JOIN query_users as query_users_mentioned ON user_mention.user_mention_id = query_users_mentioned.user_id WHERE status.created_at > '%s';",format(Sys.Date()-7))
usermentions.df <- DBI::dbGetQuery(con,usermention.query)
if(nrow(usermentions.df) > 0){
  g <- graph_from_data_frame(usermentions.df)
  vertices.df <- data.frame(index=as.integer(V(g)), screen_name = names(V(g)),stringsAsFactors = FALSE)
  vertex.query <- sprintf("SELECT query_users.screen_name,query_users.party,query_users.body,user.name,user.location,user.description,user.followers_count,user.statuses_count FROM query_users JOIN user ON query_users.user_id = user.id WHERE query_users.screen_name IN (%s);",paste("'",vertices.df$screen_name,"'",sep="",collapse=","))
  vertex.query.df <- DBI::dbGetQuery(con,vertex.query)
  vertex.query.df <- vertex.query.df[match(vertices.df$screen_name,vertex.query.df$screen_name),]

  bluered <- function(party){
    if(tolower(party)=="democrat"){
      return("blue")
    } else if(tolower(party)=="republican"){
      return("red")
    } else {
      return("black")
    }
  }

  vertex.query.df$color <- sapply(vertex.query.df$party,bluered)
}
```

#### Eigenvector Centrality

```{r}
if(nrow(usermentions.df) > 0){
  ec <- centr_eigen(g,directed=FALSE)
  df <- data.frame(
    Screen_Name = paste(
      "@",
      vertex.query.df$screen_name,
      sep=""
    ),
    Eig_Centrality = as.numeric(ec$vector)
  )
  df <- df[order(df[,2],decreasing=TRUE),]
  knitr::kable(
    df[1:10,],
    row.names = FALSE,
    caption = "User Mention Network Eigenvector Centrality"
  )
}
```

#### Betweenness centrality

```{r}
if(nrow(usermentions.df) > 0){
  bc <- centr_betw(g,directed=FALSE)
  df <- data.frame(
    Screen_Name = paste(
      "@",
      vertex.query.df$screen_name,
      sep=""
    ),
    Bet_Centrality = as.numeric(as.numeric(bc$res))
  )
  df <- df[order(df[,2],decreasing=TRUE),]
  knitr::kable(
    df[1:10,],
    caption = "User Mention Network Betweenness Centrality"
  )
}
```

#### Clustering (Omitted)

#### Visualize

```{r fig.width=7, fig.height=7}
if(nrow(usermentions.df) > 0){
  verts <- data.frame(
      name = vertex.query.df$screen_name,
      label = vertex.query.df$screen_name,
      title = vertex.query.df$name,
      color = vertex.query.df$color,
      stringsAsFactors = FALSE
  )

  for(n in names(verts)){
    g <- set_vertex_attr(g,n,value=verts[,n])
  }

  plot(
    g,
    vertex.color=vertex_attr(g,"color"),
    vertex.size=4,
    vertex.label=NA
  )
}
```

#### Interactive visualization with sigmajs

```{r eval = FALSE, fig.width=7, fig.height=7}
# Not Run
if(nrow(usermentions.df) > 0){
  s <- sigmajs()
  s <- sg_from_igraph(s,g)
  s <- sg_settings(s,defaultEdgeColor="#AAA", edgeColor="default",minNodeSize=1,maxNodeSize=4)
  s
}
```

#### Interactive visualization with visNetwork
```{r eval = FALSE, fig.width=7, fig.height=7}
# Not Run
if(nrow(usermentions.df) > 0){
  g <- set_edge_attr(g,"title",E(g),paste(edge_attr(g,"created_at"),edge_attr(g,"status_text"),sep=": "))

  v <- visIgraph(g)
  v <- visEdges(
    v,
    color="gray"
  )
  v <- visNodes(
    v,
    size = 15
  )
  v
}
```

### Retweet network analysis

This is a final network model, using the retweet network.

```{r fig.width=7, fig.height=7}
retweet.query <- sprintf("SELECT query_users_retweeter.screen_name as retweeter,query_users_retweeted.screen_name as retweeted, status_retweet.id as status_id, status_retweet.text as retweet_text, status_original.text as original_text, status_retweet.created_at as created_at FROM query_users as query_users_retweeter JOIN status as status_retweet ON query_users_retweeter.user_id = status_retweet.user_id JOIN status as status_original ON status_retweet.retweet_status_id = status_original.id JOIN query_users as query_users_retweeted ON status_original.user_id = query_users_retweeted.user_id WHERE status_retweet.created_at > '%s' AND status_retweet.retweet = 1;",format(Sys.Date()-7))
retweets.df <- DBI::dbGetQuery(con,retweet.query)
if(nrow(retweets.df) > 0){
  g <- graph_from_data_frame(retweets.df)
  vertices.df <- data.frame(index=as.integer(V(g)), screen_name = names(V(g)),stringsAsFactors = FALSE)
  vertex.query <- sprintf("SELECT query_users.screen_name,query_users.party,query_users.body,user.name,user.location,user.description,user.followers_count,user.statuses_count FROM query_users JOIN user ON query_users.user_id = user.id WHERE query_users.screen_name IN (%s);",paste("'",vertices.df$screen_name,"'",sep="",collapse=","))
  vertex.query.df <- DBI::dbGetQuery(con,vertex.query)
  vertex.query.df <- vertex.query.df[match(vertices.df$screen_name,vertex.query.df$screen_name),]

  bluered <- function(party){
    if(tolower(party)=="democrat"){
      return("blue")
    } else if(tolower(party)=="republican"){
      return("red")
    } else {
      return("black")
    }
  }
  vertex.query.df$color <- sapply(vertex.query.df$party,bluered)
}
```

#### Eigenvector Centrality

```{r}
if(nrow(retweets.df) > 0){
  ec <- centr_eigen(g,directed=FALSE)
  df <- data.frame(
    Screen_Name = paste(
      "@",
      vertex.query.df$screen_name,
      sep=""
    ),
    Eig_Centrality = as.numeric(ec$vector)
  )
  df <- df[order(df[,2],decreasing=TRUE),]
  knitr::kable(
    df[1:10,],
    caption = "Retweet Network Eigenvector Centrality"
  )
}
```

#### Betweenness Centrality

```{r}
if(nrow(retweets.df) > 0){
  bc <- centr_betw(g,directed=FALSE)
  df <- data.frame(
    Screen_Name = paste(
      "@",
      vertex.query.df$screen_name,
      sep=""
    ),
    Bet_Centrality = as.numeric(bc$res)
  )
  df <- df[order(df[,2],decreasing=TRUE),]
  knitr::kable(
    df[1:10,],
    caption = "Retweet Network Betweenness Centrality"
  )
}
```

#### Clustering (Omitted)

#### Visualize

```{r fig.width=7, fig.height=7}
if(nrow(retweets.df) > 0){
  verts <- data.frame(
      name = vertex.query.df$screen_name,
      label = vertex.query.df$screen_name,
      title = vertex.query.df$name,
      color = vertex.query.df$color,
      stringsAsFactors = FALSE
  )

  for(n in names(verts)){
    g <- set_vertex_attr(g,n,value=verts[,n])
  }

  plot(g,vertex.color=vertex_attr(g,"color"),vertex.size=4,vertex.label=NA)
}
```

#### Interactive visualization with sigmajs

```{r eval = FALSE, fig.width=7, fig.height=7}
# Not Run
if(nrow(retweets.df) > 0){
  s <- sigmajs()
  s <- sg_from_igraph(s,g)
  s <- sg_settings(s,defaultEdgeColor="#AAA", edgeColor="default",minNodeSize=1,maxNodeSize=4)
  s
}
```

#### Interactive visualization with visNetwork
```{r eval = FALSE, fig.width=7, fig.height=7}
# Not Run
if(nrow(retweets.df) > 0){
  g <- set_edge_attr(g,"title",E(g),paste(edge_attr(g,"created_at"),edge_attr(g,"retweet_text"),sep=": "))

  v <- visIgraph(g)
  v <- visEdges(
    v,
    color="gray"
  )
  v <- visNodes(
    v,
    size = 15
  )
  v
}
```

# Way ahead

There are many more interesting analyses that could be done with this set.  For example, do how to democrats' and republicans' sentiments vary when they use different hashtags or user mentions?  Can we classify users as democrat or republican by their Twitter timelines?  Hopefully, this package provides a useful toolset in collecting and analyzing Twitter data to answer questions like these.


## Don't forget to disconnect!

```{r}
DBI::dbDisconnect(con)
```









